### 2023-03-29 response-to-audit-report.md

     , Spec.Marlowe.Plutus.ScriptContext.tests
-    , Spec.Marlowe.Plutus.Specification.tests
+    , Spec.Marlowe.Plutus.Specification.tests referencePaths
     ]
-
diff --git a/marlowe-test/src/Spec/Marlowe/Plutus/MList.hs b/marlowe-test/src/Spec/Marlowe/Plutus/MList.hs
new file mode 100644
index 000000000..7edb3424e
--- /dev/null
+++ b/marlowe-test/src/Spec/Marlowe/Plutus/MList.hs
@@ -0,0 +1,312 @@
+-----------------------------------------------------------------------------
+--
+-- Module      :  $Headers
+-- License     :  Apache 2.0
+--
+-- Stability   :  Experimental
+-- Portability :  Portable
+--
+-- | Compare the `PlutusTx.AssocMap` to Isabelle's `MList`.
+--
+-----------------------------------------------------------------------------
+
+
+module Spec.Marlowe.Plutus.MList
+  ( -- * Testing
+    tests
+  ) where
+
+
+import Data.Function (on)
+import Data.List (nubBy, sortBy)
+import Data.Maybe (fromMaybe, isJust)
+import Prelude hiding (lookup)
+import Test.Tasty (TestTree, testGroup)
+import Test.Tasty.HUnit (Assertion, assertBool, testCase)
+import Test.Tasty.QuickCheck (Arbitrary(..), Gen, Property, elements, forAll, property, testProperty)
+
+import qualified PlutusTx.AssocMap as AM
+  (Map, delete, empty, fromList, insert, lookup, member, null, singleton, toList, unionWith)
+import qualified PlutusTx.Eq as P (Eq)
+
+
+-- | An association list in Isabelle.
+type MList a b = [(a, b)]
+
+
+-- | Empty Isabelle `MList`.
+--
+-- @
+-- definition empty :: "('a::linorder \<times> 'b) list" where
+--   "empty = Nil"
+-- @
+empty :: MList a b
+empty = []
+
+
+{-# ANN insert "HLint: ignore Use guards" #-}
+{-# ANN insert "HLint: ignore Use list literal" #-}
+
+-- | Insert into Isabelle `MList`.
+--
+-- @
+-- fun insert :: "'a::linorder \<Rightarrow> 'b \<Rightarrow> ('a \<times> 'b) list \<Rightarrow> ('a \<times> 'b) list" where
+--   "insert a b Nil = Cons (a, b) Nil" |
+--   "insert a b (Cons (x, y) z) =
+--     (if a < x
+--      then (Cons (a, b) (Cons (x, y) z))
+--      else (if a > x
+--            then (Cons (x, y) (insert a b z))
+--            else (Cons (x, b) z)))"
+-- @
+insert :: Ord a => a -> b -> MList a b -> MList a b
+insert a b [] = (a, b) : []
+insert a b ((x, y) : z) =
+  if a < x
+    then (a, b) : (x, y) : z
+    else if a > x
+      then (x, y) : insert a b z
+      else (x, b) : z
+
+
+{-# ANN delete "HLint: ignore Use guards" #-}
+
+-- | Delete from Isabelle `MList`.
+--
+-- @
+-- fun delete :: "'a::linorder \<Rightarrow> ('a \<times> 'b) list \<Rightarrow> ('a \<times> 'b) list" where
+--   "delete a Nil = Nil" |
+--   "delete a (Cons (x, y) z) =
+--     (if a = x
+--      then z
+--      else (if a > x
+--            then (Cons (x, y) (delete a z))
+--            else (Cons (x, y) z)))"
+-- @
+delete :: Ord a => a -> MList a b -> MList a b
+delete _a [] = []
+delete a ((x, y) : z) =
+  if a == x
+    then z
+    else if a > x
+      then (x, y) : delete a z
+      else (x, y) : z
+
+
+{-# ANN lookup "HLint: ignore Use guards" #-}
+
+-- | Lookup in Isabelle `MList`.
+--
+-- @
+-- fun lookup :: "'a::linorder \<Rightarrow> ('a \<times> 'b) list \<Rightarrow> 'b option" where
+--   "lookup a Nil = None" |
+--   "lookup a (Cons (x, y) z) =
+--     (if a = x
+--      then Some y
+--      else (if a > x
+--            then lookup a z
+--            else None))"
+-- @
+lookup :: Ord a => a -> MList a b -> Maybe b
+lookup _a [] = Nothing
+lookup a ((x, y) : z) =
+  if a == x
+    then Just y
+    else if a > x
+      then lookup a z
+      else Nothing
+
+
+{-# ANN unionWith "HLint: ignore Use guards" #-}
+
+-- | Union with Isabelle `MList`.
+--
+-- @
+-- fun unionWith :: "('b \<Rightarrow> 'b \<Rightarrow> 'b) \<Rightarrow> ('a \<times> 'b) list \<Rightarrow>
+--                   ('a \<times> 'b) list \<Rightarrow> (('a::linorder) \<times> 'b) list" where
+--   "unionWith f (Cons (x, y) t) (Cons (x2, y2) t2) =
+--     (if x < x2
+--      then Cons (x, y) (unionWith f t (Cons (x2, y2) t2))
+--      else (if x > x2
+--            then Cons (x2, y2) (unionWith f (Cons (x, y) t) t2)
+--            else Cons (x, f y y2) (unionWith f t t2)))" |
+--   "unionWith f Nil l = l" |
+--   "unionWith f l Nil = l"
+-- @
+unionWith :: Ord a => (b -> b -> b) -> MList a b -> MList a b -> MList a b
+unionWith f ((x, y) : t) ((x2, y2) : t2) =
+  if x < x2
+    then (x, y) : unionWith f t ((x2, y2) : t2)
+    else if x > x2
+      then (x2, y2) : unionWith f ((x, y) : t) t2
+      else (x, f y y2) : unionWith f t t2
+unionWith _f [] l = l
+unionWith _f l [] = l
+
+
+-- | Find with default for Isabelle `MList`.
+--
+-- @
+-- fun findWithDefault :: "'b \<Rightarrow> 'a \<Rightarrow> (('a::linorder) \<times> 'b) list \<Rightarrow> 'b" where
+--   "findWithDefault d k l = (case lookup k l of
+--                               None \<Rightarrow> d
+--                             | Some x \<Rightarrow> x)"
+-- @
+findWithDefault :: Ord a => b -> a -> MList a b -> b
+findWithDefault d k l =
+  case lookup k l of
+    Nothing -> d
+    Just x -> x
+
+
+-- | Run tests.
+tests :: TestTree
+tests =
+  testGroup "MList vs AssocMap"
+    [
+      testCase     "`empty` is equivalent"           checkEmpty
+    , testProperty "`null` is equivalent"            checkNull
+    , testProperty "`singleton` is equivalent"       checkSingleton
+    , testProperty "`insert` is equivalent"          checkInsert
+    , testProperty "`delete` is equivalent"          checkDelete
+    , testProperty "`lookup` is equivalent"          checkLookup
+    , testProperty "`member` is equivalent"          checkMember
+    , testProperty "`unionWith` is equivalent"       checkUnionWith
+    , testProperty "`findWithDefault` is equivalent" checkFindWithDefault
+    ]
+
+
+-- | Key type for testing.
+type Key = Integer
+
+
+-- | Value type for testing.
+type Value = [()]
+
+
+-- | Generate a sorted `MList` with no duplicate keys.
+arbitraryMList :: Gen (MList Key Value, AM.Map Key Value)
+arbitraryMList =
+  do
+    items <- nubBy ((==) `on` fst) <$> arbitrary
+    pure (sortBy (compare `on` fst) items, AM.fromList items)
+
+
+-- | Compare an `MList` to an `AssocMap`, ignoring ordering.
+equivalent :: Ord a => P.Eq a => Eq b => MList a b -> AM.Map a b -> Bool
+equivalent mlist assocmap =
+  mlist == sortBy (compare `on` fst) (AM.toList assocmap)
+    && and [AM.lookup a assocmap == Just b | (a, b) <- mlist]
+    && and [lookup a mlist == Just b | (a, b) <- AM.toList assocmap]
+
+
+-- | Compare `empty` for `MList` and `AssocMap`, provided the `MList` is sorted and neither contains duplicate keys.
+checkEmpty :: Assertion
+checkEmpty = assertBool "Empty MList and AssocMap" $ (empty :: MList Key Value) `equivalent` AM.empty
+
+
+-- | Compare `null` for `MList` and `AssocMap`, provided the `MList` is sorted and neither contains duplicate keys.
+checkNull :: Property
+checkNull = property $ do
+  let
+    gen = do
+      (mlist, assocmap) <- arbitraryMList
+      pure (mlist, assocmap)
+  forAll gen
+    $ \(mlist, assocmap) -> (== empty) mlist == AM.null assocmap
+
+
+-- | Compare `singleton` for `MList` and `AssocMap`, provided the `MList` is sorted and neither contains duplicate keys.
+checkSingleton :: Property
+checkSingleton = property $ do
+  let
+    gen = do
+      a <- arbitrary :: Gen Key
+      b <- arbitrary :: Gen Value
+      pure (a, b)
+  forAll gen
+    $ \(a, b) -> [(a, b)] `equivalent` AM.singleton a b
+
+
+-- | Compare `insert` for `MList` and `AssocMap`, provided the `MList` is sorted and neither contains duplicate keys.
+checkInsert :: Property
+checkInsert = property $ do
+  let
+    gen = do
+      (mlist, assocmap) <- arbitraryMList
+      a <- arbitrary
+      b <- arbitrary
+      pure (mlist, assocmap, a, b)
+  forAll gen
+    $ \(mlist, assocmap, a, b) -> insert a b mlist `equivalent` AM.insert a b assocmap
+
+
+-- | Compare `delete` for `MList` and `AssocMap`, provided the `MList` is sorted and neither contains duplicate keys.
+checkDelete :: Property
+checkDelete = property $ do
+  let
+    gen = do
+      (mlist, assocmap) <- arbitraryMList
+      a <- arbitrary
+      pure (mlist, assocmap, a)
+  forAll gen
+    $ \(mlist, assocmap, a) -> delete a mlist `equivalent` AM.delete a assocmap
+
+
+-- | Compare `lookup` for `MList` and `AssocMap`, provided the `MList` is sorted and neither contains duplicate keys.
+checkLookup :: Property
+checkLookup = property $ do
+  let
+    gen = do
+      (mlist, assocmap) <- arbitraryMList
+      a <- arbitrary
+      pure (mlist, assocmap, a)
+  forAll gen
+    $ \(mlist, assocmap, a) -> lookup a mlist == AM.lookup a assocmap
+
+
+-- | Compare `member` for `MList` and `AssocMap`, provided the `MList` is sorted and neither contains duplicate keys.
+checkMember :: Property
+checkMember = property $ do
+  let
+    gen = do
+      (mlist, assocmap) <- arbitraryMList
+      a <- arbitrary
+      pure (mlist, assocmap, a)
+  forAll gen
+    $ \(mlist, assocmap, a) -> isJust (lookup a mlist) == AM.member a assocmap
+
+
+-- | Compare `unionWith` for `MList` and `AssocMap`, provided the `MList` is sorted and neither contains duplicate keys.
+checkUnionWith :: Property
+checkUnionWith = property $ do
+  let
+    gen = do
+      (mlist, assocmap) <- arbitraryMList
+      (mlist', assocmap') <- arbitraryMList
+      function <- elements ["concat", "shortest", "longest", "first", "second"]
+      pure (mlist, assocmap, mlist', assocmap', function)
+  forAll gen
+    $ \(mlist, assocmap, mlist', assocmap', function) ->
+      let
+        f = case function of
+              "shortest" -> \x y -> if length x < length y then x else y
+              "longest"  -> \x y -> if length x > length y then x else y
+              "first"    -> const
+              "second"   -> const id
+              _          -> (<>)
+      in
+        unionWith f mlist mlist' `equivalent` AM.unionWith f assocmap assocmap'
+
+
+-- | Compare `findWithDefault` for `MList` and `AssocMap`, provided the `MList` is sorted and neither contains duplicate keys.
+checkFindWithDefault :: Property
+checkFindWithDefault = property $ do
+  let
+    gen = do
+      (mlist, assocmap) <- arbitraryMList
+      a <- arbitrary
+      b <- arbitrary
+      pure (mlist, assocmap, a, b)
+  forAll gen
+    $ \(mlist, assocmap, a, b) -> findWithDefault b a mlist == fromMaybe b (AM.lookup a assocmap)
diff --git a/marlowe-test/src/Spec/Marlowe/Plutus/Script.hs b/marlowe-test/src/Spec/Marlowe/Plutus/Script.hs
index ccfab1c9a..9ced8d81e 100644
--- a/marlowe-test/src/Spec/Marlowe/Plutus/Script.hs
+++ b/marlowe-test/src/Spec/Marlowe/Plutus/Script.hs
@@ -11,7 +11,9 @@
 -----------------------------------------------------------------------------

+{-# LANGUAGE NumericUnderscores #-}
 {-# LANGUAGE OverloadedStrings #-}
+{-# LANGUAGE RecordWildCards #-}

 module Spec.Marlowe.Plutus.Script
@@ -53,6 +55,7 @@ import Plutus.V2.Ledger.Api
   , Data
   , Datum(Datum)
   , DatumHash
+  , ExBudget(..)
   , ToData(toBuiltinData)
   , TokenName
   , Validator(getValidator)
@@ -64,6 +67,11 @@ import qualified Data.ByteString.Short as SBS (ShortByteString, toShort)
 import qualified Data.Map.Strict as M (fromList)

+-- | Check the Plutus execution budget.
+enforceBudget :: Bool
+enforceBudget = False
+
+
 -- | Run the Plutus evaluator on the Marlowe semantics validator.
 evaluateSemantics :: Data                    -- ^ The datum.
                   -> Data                    -- ^ The redeemer.
@@ -72,9 +80,11 @@ evaluateSemantics :: Data                    -- ^ The datum.
 evaluateSemantics datum redeemer context =
   case evaluationContext of
     Left message -> This message
-    Right ec     -> case evaluateScriptCounting PlutusV2 (ProtocolVersion 7 0) Verbose ec serialiseSemanticsValidator [datum, redeemer, context] of
-                      (logOutput, Right _     ) -> That logOutput
-                      (logOutput, Left message) -> These (show message) logOutput
+    Right ec     -> case evaluateScriptCounting PlutusV2 (ProtocolVersion 8 0) Verbose ec serialiseSemanticsValidator [datum, redeemer, context] of
+                      (logOutput, Right ex@ExBudget{..}) -> if enforceBudget && (exBudgetCPU > 10_000_000_000 || exBudgetMemory > 14_000_000)
+                                                              then These ("Exceeded Plutus budget: " <> show ex) logOutput
+                                                              else That logOutput
+                      (logOutput, Left message         ) -> These (show message) logOutput

 -- | Run the Plutus evaluator on the Marlowe payout validator.
@@ -321,12 +331,12 @@ costModel =
     , ("unListData-memory-arguments", 32)
     , ("unMapData-cpu-arguments", 38314)
     , ("unMapData-memory-arguments", 32)
-    , ("verifyEcdsaSecp256k1Signature-cpu-arguments", 20000000000)
-    , ("verifyEcdsaSecp256k1Signature-memory-arguments", 20000000000)
-    , ("verifyEd25519Signature-cpu-arguments-intercept", 9462713)
-    , ("verifyEd25519Signature-cpu-arguments-slope", 1021)
+    , ("verifyEcdsaSecp256k1Signature-cpu-arguments", 35892428)
+    , ("verifyEcdsaSecp256k1Signature-memory-arguments", 10)
+    , ("verifyEd25519Signature-cpu-arguments-intercept", 57996947)
+    , ("verifyEd25519Signature-cpu-arguments-slope", 18975)
     , ("verifyEd25519Signature-memory-arguments", 10)
-    , ("verifySchnorrSecp256k1Signature-cpu-arguments-intercept", 20000000000)
-    , ("verifySchnorrSecp256k1Signature-cpu-arguments-slope", 0)
-    , ("verifySchnorrSecp256k1Signature-memory-arguments", 20000000000)
+    , ("verifySchnorrSecp256k1Signature-cpu-arguments-intercept", 38887044)
+    , ("verifySchnorrSecp256k1Signature-cpu-arguments-slope", 32947)
+    , ("verifySchnorrSecp256k1Signature-memory-arguments", 10)
     ]
diff --git a/marlowe-test/src/Spec/Marlowe/Plutus/Specification.hs b/marlowe-test/src/Spec/Marlowe/Plutus/Specification.hs
index e02c42079..fc8d92b8f 100644
--- a/marlowe-test/src/Spec/Marlowe/Plutus/Specification.hs
+++ b/marlowe-test/src/Spec/Marlowe/Plutus/Specification.hs
@@ -23,9 +23,10 @@ module Spec.Marlowe.Plutus.Specification
   ) where

### 2023-03-29 response-to-audit-report.md

-import Control.Lens (use, uses, (%=), (<>=), (^.))
+import Control.Lens (use, uses, (%=), (<>=), (<~), (^.))
 import Control.Monad.State (lift)
 import Data.Bifunctor (bimap)
+import Data.List (nub)
 import Data.Maybe (maybeToList)
 import Data.Proxy (Proxy(..))
 import Data.These (These(That, These, This))
@@ -35,6 +36,7 @@ import Language.Marlowe.Core.V1.Semantics
   , Payment(Payment)
   , TransactionInput(..)
   , TransactionOutput(txOutContract, txOutPayments, txOutState)
+  , totalBalance
   )
 import Language.Marlowe.Core.V1.Semantics.Types
   ( ChoiceId(ChoiceId)
@@ -54,7 +56,7 @@ import Plutus.V2.Ledger.Api
   , BuiltinData(BuiltinData)
   , Credential(PubKeyCredential)
   , Data(B, Constr, List)
-  , Datum(Datum)
+  , Datum(..)
   , DatumHash(DatumHash)
   , FromData(..)
   , OutputDatum(..)
@@ -71,12 +73,14 @@ import Plutus.V2.Ledger.Api
   , singleton
   , toData
   )
+import Spec.Marlowe.Plutus.Lens ((<><~))
 import Spec.Marlowe.Plutus.Script
   (evaluatePayout, evaluateSemantics, payoutAddress, payoutScriptHash, semanticsAddress, semanticsScriptHash)
 import Spec.Marlowe.Plutus.Transaction
   ( ArbitraryTransaction
   , arbitraryPayoutTransaction
   , arbitrarySemanticsTransaction
+  , isScriptTxIn
   , merkleize
   , noModify
   , noVeto
@@ -97,26 +101,41 @@ import Spec.Marlowe.Plutus.Types
   , output
   , role
   )
+import Spec.Marlowe.Reference (ReferencePath)
 import Spec.Marlowe.Semantics.Arbitrary (arbitraryPositiveInteger)
 import Test.Tasty (TestTree, testGroup)
-import Test.Tasty.QuickCheck (Arbitrary(..), Gen, Property, forAll, property, suchThat, testProperty, (===))
+import Test.Tasty.QuickCheck
+  ( Arbitrary(..)
+  , Gen
+  , Property
+  , chooseInteger
+  , elements
+  , forAll
+  , listOf1
+  , oneof
+  , property
+  , suchThat
+  , testProperty
+  , (===)
+  )

 import qualified Language.Marlowe.Core.V1.Semantics as M (MarloweData(marloweParams))
-import qualified Language.Marlowe.Core.V1.Semantics.Types as M (Party(Address))
-import qualified PlutusTx.AssocMap as AM (fromList, insert, toList)
+import qualified Language.Marlowe.Core.V1.Semantics.Types as M (Party(Address), State(..))
+import qualified PlutusTx.AssocMap as AM (Map, fromList, insert, keys, null, toList)
+import qualified Test.Tasty.QuickCheck as Q (shuffle)

 -- | Run tests.
-tests :: TestTree
-tests =
+tests :: [ReferencePath] -> TestTree
+tests referencePaths =
   testGroup "Marlowe On-Chain Specification"
     [
       testGroup "Semantics Validator"
         [
           testGroup "Valid transaction succeeds"
             [
-              testProperty "Noiseless" $ checkSemanticsTransaction noModify noModify noVeto True False
-            , testProperty "Noisy"     $ checkSemanticsTransaction noModify noModify noVeto True True
+              testProperty "Noiseless" $ checkSemanticsTransaction referencePaths noModify noModify noVeto True False True
+            , testProperty "Noisy"     $ checkSemanticsTransaction referencePaths noModify noModify noVeto True True  True
             ]
         , testGroup "Constraint 1. Typed validation"
             [
@@ -129,23 +148,23 @@ tests =
             ]
         , testGroup "Constraint 2. Single Marlowe script input"
             [
-              testProperty "Invalid attempt to run two Marlowe scripts" checkDoubleInput
+              testProperty "Invalid attempt to run two Marlowe scripts" $ checkDoubleInput referencePaths
             ]
         , testGroup "Constraint 3. Single Marlowe output"
             [
-              testProperty "Invalid attempt to split Marlowe output" checkMultipleOutput
+              testProperty "Invalid attempt to split Marlowe output" $ checkMultipleOutput referencePaths
             ]
         , testGroup "Constraint 4. No output to script on close"
             [
-              testProperty "Invalid attempt to output to Marlowe on close" checkCloseOutput
+              testProperty "Invalid attempt to output to Marlowe on close" $ checkCloseOutput referencePaths
             ]
         , testGroup "Constraint 5. Input value from script"
             [
-              testProperty "Invalid mismatch between state and script input" checkValueInput
+              testProperty "Invalid mismatch between state and script input" $ checkValueInput referencePaths
             ]
         , testGroup "Constraint 6. Output value to script"
             [
-              testProperty "Invalid mismatch between state and script input" checkValueOutput
+              testProperty "Invalid mismatch between expected and actual output to script" $ checkValueOutput referencePaths
             ]
         , testGroup "Constraint 7. Input state"
             [
@@ -157,39 +176,53 @@ tests =
             ]
         , testGroup "Constraint 9. Marlowe parameters"
             [
-              testProperty "Invalid alteration of parameters." checkParamsOutput
+              testProperty "Invalid alteration of parameters" $ checkParamsOutput referencePaths
             ]
         , testGroup "Constraint 10. Output state"
             [
-              testProperty "Invalid mismatch between state and script output" checkStateOutput
+              testProperty "Invalid mismatch between state and script output's state" $ checkStateOutput referencePaths
             ]
         , testGroup "Constraint 11. Output contract"
             [
-              testProperty "Invalid mismatch between contract and script output" checkContractOutput
+              testProperty "Invalid mismatch between contract and script output" $ checkContractOutput referencePaths
             ]
         , testGroup "Constraint 12. Merkleized continuations"
             [
-              testProperty "Valid merkleization"   $ checkMerkleization True
-            , testProperty "Invalid merkleization" $ checkMerkleization False
+              testProperty "Valid merkleization"   $ checkMerkleization referencePaths True
+            , testProperty "Invalid merkleization" $ checkMerkleization referencePaths False
             ]
         , testGroup "Constraint 13. Positive balances"
             [
-              testProperty "Invalid non-positive balance" checkPositiveAccounts
+              testProperty "Invalid non-positive balance" $ checkPositiveAccounts referencePaths
+              -- TODO: This test on the output state requires instrumenting the Plutus script. For now, this constraint is enforced manually by code inspection.
             ]
         , testGroup "Constraint 14. Inputs authorized"
             [
-              testProperty "Invalid missing authorization" checkAuthorization
+              testProperty "Invalid missing authorization" $ checkAuthorization referencePaths
             ]
         , testGroup "Constraint 15. Sufficient payment"
             [
-              testProperty "Invalid insufficient payment" checkPayment
+              testProperty "Invalid insufficient payment" $ checkPayment referencePaths
+            ]
+        , testGroup "Constraint 18. Final balance"
+            [
+              testProperty "Invalid mismatch between output value and state" $ checkOutputConsistency referencePaths
+            ]
+        , testGroup "Constraint 19. No duplicates"
+            [
+              testProperty "Invalid duplicate accounts in input state" $ checkInputDuplicates referencePaths
+              -- TODO: This test on the output state requires instrumenting the Plutus script. For now, this constraint is enforced manually by code inspection.
+            ]
+        , testGroup "Constraint 20. Single satisfaction"
+            [
+              testProperty "Invalid other validators during payment" $ checkOtherValidators referencePaths
             ]
         , testProperty "Script hash matches reference hash"
             $ checkValidatorHash semanticsScriptHash
               -- DO NOT ALTER THE FOLLOWING VALUE UNLESS YOU ARE COMMITTING
               -- APPROVED CHANGES TO MARLOWE'S SEMANTICS VALIDATOR. THIS HASH
               -- HAS IMPLICATIONS FOR VERSIONING, AUDIT, AND CONTRACT DISCOVERY.
-              "6a9391d6aa51af28dd876ebb5565b69d1e83e5ac7861506bd29b56b0"
+              "2ed2631dbb277c84334453c5c437b86325d371f0835a28b910a91a6e"
         ]
     , testGroup "Payout Validator"
         [
@@ -217,7 +250,7 @@ tests =
               -- DO NOT ALTER THE FOLLOWING VALUE UNLESS YOU ARE COMMITTING
               -- APPROVED CHANGES TO MARLOWE'S ROLE VALIDATOR. THIS HASH HAS
               -- IMPLICATIONS FOR VERSIONING, AUDIT, AND CONTRACT DISCOVERY.
-              "49076eab20243dc9462511fb98a9cfb719f86e9692288139b7c91df3"
+              "e165610232235bbbbeff5b998b233daae42979dec92a6722d9cda989"
         ]
     ]

@@ -260,19 +293,21 @@ check1Invalid _ allowEmptyList allowByteString allowUnit =

 -- | Check that a semantics transaction succeeds.
-checkSemanticsTransaction :: ArbitraryTransaction SemanticsTransaction ()      -- ^ Modifications to make before building the valid transaction.
+checkSemanticsTransaction :: [ReferencePath]                                   -- ^ The reference execution paths from which to choose.
+                          -> ArbitraryTransaction SemanticsTransaction ()      -- ^ Modifications to make before building the valid transaction.
                           -> ArbitraryTransaction SemanticsTransaction ()      -- ^ Modifications to make after building the valid transaction.
                           -> (PlutusTransaction SemanticsTransaction -> Bool)  -- ^ Whether to discard the transaction from the testing.
                           -> Bool                                              -- ^ Whether the transaction should test as valid.
                           -> Bool                                              -- ^ Whether to add noise to the script context.
+                          -> Bool                                              -- ^ Whether to allow merkleization.
                           -> Property                                          -- ^ The test property.
-checkSemanticsTransaction modifyBefore modifyAfter condition valid noisy =
+checkSemanticsTransaction referencePaths modifyBefore modifyAfter condition valid noisy allowMerkleization =
   property
-    . forAll (arbitrarySemanticsTransaction modifyBefore modifyAfter noisy `suchThat` condition)
+    . forAll (arbitrarySemanticsTransaction referencePaths modifyBefore modifyAfter noisy allowMerkleization `suchThat` condition)
     $ \PlutusTransaction{..} ->
       case evaluateSemantics (toData _datum) (toData _redeemer) (toData _scriptContext) of
-        This  e   -> not valid || (error $ show e)
-        These e l -> not valid || (error $ show e <> ": " <> show l)
+        This  e   -> not valid || error (show e)
+        These e l -> not valid || error (show e <> ": " <> show l)
         That    _ -> valid

@@ -288,14 +323,14 @@ checkPayoutTransaction modifyBefore modifyAfter condition valid noisy =
     . forAll (arbitraryPayoutTransaction modifyBefore modifyAfter noisy `suchThat` condition)
     $ \PlutusTransaction{..} ->
       case evaluatePayout (toData _datum) (toData _redeemer) (toData _scriptContext) of
-        This  e   -> not valid || (error $ show e)
-        These e l -> not valid || (error $ show e <> ": " <> show l)
+        This  e   -> not valid || error (show e)
+        These e l -> not valid || error (show e <> ": " <> show l)
         That    _ -> valid

 -- | Check that validation fails if two Marlowe scripts are run.
-checkDoubleInput :: Property
-checkDoubleInput =
+checkDoubleInput :: [ReferencePath] -> Property
+checkDoubleInput referencePaths =
   let
     modifyAfter =
       do
@@ -313,7 +348,7 @@ checkDoubleInput =
         infoData <>= AM.fromList [(inDatumHash, inDatum)]
         shuffle
   in
-    checkSemanticsTransaction noModify modifyAfter noVeto False False
+    checkSemanticsTransaction referencePaths noModify modifyAfter noVeto False False False

 -- | Split a value in half.
@@ -341,9 +376,19 @@ notCloses :: PlutusTransaction SemanticsTransaction -> Bool
 notCloses = not . doesClose

+-- | Ensure that the contract makes payments.
+hasPayouts :: PlutusTransaction SemanticsTransaction -> Bool
+hasPayouts =
+  let
+    isPayout (Payment _ (Party _) _ i) = i > 0
+    isPayout _ = False
+  in
+    any isPayout . txOutPayments . (^. output)
+
+
 -- | Check that validation fails if there is more than one Marlowe output.
-checkMultipleOutput :: Property
-checkMultipleOutput =
+checkMultipleOutput :: [ReferencePath] -> Property
+checkMultipleOutput referencePaths =
   let
     modifyAfter =
       do
@@ -356,12 +401,12 @@ checkMultipleOutput =
         infoOutputs %= concatMap splitOwnOutput
         shuffle
   in
-    checkSemanticsTransaction noModify modifyAfter notCloses False False
+    checkSemanticsTransaction referencePaths noModify modifyAfter notCloses False False False

 -- | Check that validation fails if there is one Marlowe output upon close.
-checkCloseOutput :: Property
-checkCloseOutput =
+checkCloseOutput :: [ReferencePath] -> Property
+checkCloseOutput referencePaths =
   let
     modifyAfter =
       do
@@ -374,12 +419,12 @@ checkCloseOutput =
         infoOutputs <>= (txInInfoResolved <$> inScript)
         shuffle
   in
-    checkSemanticsTransaction noModify modifyAfter doesClose False False
+    checkSemanticsTransaction referencePaths noModify modifyAfter doesClose False False False

 -- | Check that value input to a script matches its input state.
-checkValueInput :: Property
-checkValueInput =
+checkValueInput :: [ReferencePath] -> Property
+checkValueInput referencePaths =
   let
     modifyAfter =
       do
@@ -391,29 +436,90 @@ checkValueInput =
         -- Update the inputs with the incremented script input.
         infoInputs %= fmap incrementOwnInput
     in
-      checkSemanticsTransaction noModify modifyAfter noVeto False False
+      checkSemanticsTransaction referencePaths noModify modifyAfter noVeto False False False

--- | Check that value output to a script matches its output state.
-checkValueOutput :: Property
-checkValueOutput =
+-- | Check that value output to a script matches its expectation.
+checkValueOutput :: [ReferencePath] -> Property
+checkValueOutput referencePaths =
   let
     modifyAfter =
       do
+        delta <- lift $ oneof [chooseInteger (-5, -1), chooseInteger (1, 5), arbitrary `suchThat` (/= 0)]  -- Ensure small non-zero integers.
         let
-          -- Add one lovelace to the output to the script.
+          -- Add or subtract some lovelace to the output to the script.
           incrementOwnOutput txOut@(TxOut address value _ _)
-            | address == semanticsAddress = txOut {txOutValue = value <> singleton adaSymbol adaToken 1}
-            | otherwise                  = txOut
+            | address == semanticsAddress = txOut {txOutValue = value <> singleton adaSymbol adaToken delta}
+            | otherwise                   = txOut
         -- Update the outputs with the incremented script output.
         infoOutputs %= fmap incrementOwnOutput
   in
-    checkSemanticsTransaction noModify modifyAfter notCloses False False
+    checkSemanticsTransaction referencePaths noModify modifyAfter notCloses False False False
+
+
+-- | Check the consistency of the output value with the output state.
+checkOutputConsistency :: [ReferencePath] -> Property
+checkOutputConsistency referencePaths =
+  property
+    . forAll (arbitrarySemanticsTransaction referencePaths noModify noModify False True)
+    $ \tx ->
+      let
+        findOwnOutput (TxOut address value _ _)
+          | address == semanticsAddress = value
+          | otherwise                   = mempty
+        outValue = foldMap findOwnOutput $ tx ^. infoOutputs
+        finalBalance = totalBalance . accounts . txOutState $ tx ^. output
+        valid = outValue == finalBalance
+      in
+        checkSemanticsTransaction referencePaths noModify noModify notCloses valid False False
+
+
+-- | Add a duplicate entry to an assocation list.
+addDuplicate :: Arbitrary v => AM.Map k v -> Gen (AM.Map k v)
+addDuplicate am =
+  do
+    let
+      am' = AM.toList am
+    key <- elements $ fst <$> am'
+    value <- arbitrary
+    AM.fromList <$> Q.shuffle ((key, value) : am')
+
+
+-- | Check for the detection of duplicates in input state
+checkInputDuplicates :: [ReferencePath] -> Property
+checkInputDuplicates referencePaths =
+  let
+    hasDuplicates tx =
+      let
+        hasDuplicate am = length (AM.keys am) /= length (nub $ AM.keys am)
+        M.State{..} = tx ^. inputState
+      in
+           hasDuplicate accounts
+        || hasDuplicate choices
+        || hasDuplicate boundValues
+    makeDuplicates am =
+      if AM.null am
+        then pure am
+        else oneof [pure am, addDuplicate am]
+    modifyBefore =
+      do
+        M.State{..} <- use inputState
+        inputState <~
+          lift
+            (
+              M.State
+                <$> makeDuplicates accounts
+                <*> makeDuplicates choices
+                <*> makeDuplicates boundValues
+                <*> pure minTime
+            )
+  in
+    checkSemanticsTransaction referencePaths modifyBefore noModify hasDuplicates False False False

 -- | Check that output datum to a script matches its semantic output.
-checkDatumOutput :: (MarloweData -> Gen MarloweData) -> Property
-checkDatumOutput perturb =
+checkDatumOutput :: [ReferencePath] -> (MarloweData -> Gen MarloweData) -> Property
+checkDatumOutput referencePaths perturb =
   let
     modifyAfter =
       do
@@ -432,13 +538,24 @@ checkDatumOutput perturb =
         -- Update the data with the modification.
         infoData %= AM.fromList . fmap perturbOwnOutputDatum . AM.toList
   in
-    checkSemanticsTransaction noModify modifyAfter notCloses False False
+    checkSemanticsTransaction referencePaths noModify modifyAfter notCloses False False False
+
+
+-- | Check that other validators are forbidden during payments.
+checkOtherValidators :: [ReferencePath] -> Property
+checkOtherValidators referencePaths =
+  let
+    modifyAfter =
+      -- Add an extra script input.
+      infoInputs <><~ lift (listOf1 $ arbitrary `suchThat` isScriptTxIn)
+  in
+    checkSemanticsTransaction referencePaths noModify modifyAfter hasPayouts False False False
